Architectures for workflow integration in science gateways

### 2016-04-12 paper.tex
Wrote section 4.7.

@@

@@

\emph{Modularity} measures the ability to replace elements in the
architecture, or to integrate new elements. Elements include
workflow engines, workflows and infrastructure. It breaks down in 4
metrics:
\begin{itemize}[itemsep=0cm]
\item New engine type: number of interactions or components to modify
  to integrate a new workflow engine type in the
  architecture. Modification of an interaction is deemed necessary
  when the parameters involved in this interaction are
  modified. Modification of a component is required when its code
  needs to be modified or recompiled (science gateway or workflow
  service), or when a new piece of software has to be installed
  (infrastructure only).
\item New engine version: number of interactions or components to
  modify to integrate a new version of a workflow engine in the
  architecture, assuming that another version of the same engine type
  is already available. We assume that different versions of a
  workflow engine share the same interface, i.e. they can be invoked
  using the same software. When this is not the case, the different
  versions have to be considered as different engine types.
\item New infrastructure: number of interactions or components to
  modify to integrate a new type of infrastructure in the
  architecture.
\item New workflow: number of interactions required to integrate a new
  workflow in the architecture.
\item Meta-workflow: ability to describe meta-workflows from existing
  workflows (1: not available; 0: available).
\end{itemize}

@@

\subsection{Nested workflows and service invocation}

@@

\begin{table*}
\centering
\begin{tabular}{rcccccc}
\textbf{Metric}                      & \textbf{Tight}
                                     & \textbf{Service}
                                     & \textbf{Sub-task}
                                     & \textbf{Pool}
                                     & \textbf{Nested}
                                     & \textbf{Import} \\
\multicolumn{7}{c}{\cellcolor[HTML]{EEEEEE}\textbf{Integration}}\\
Science gateway -- \texttt{I$_1$}    & \cellcolor[HTML]{99FF99}1
                                     & \cellcolor[HTML]{99FF99}1
                                     & \cellcolor[HTML]{99FF99}0  
                                     & \cellcolor[HTML]{99FF99}1
                                     & \cellcolor[HTML]{99FF99}1
                                     & \cellcolor[HTML]{99AA99}3 \\
Workflow engine -- \texttt{I$_2$}    & \cellcolor[HTML]{99AA99}2
                                     & \cellcolor[HTML]{99FF99}3 
                                     & \cellcolor[HTML]{99FF99}1
                                     & \cellcolor[HTML]{99AA99}2
                                     & \cellcolor[HTML]{99AA99}5
                                     & \cellcolor[HTML]{99FF99}1 \\
Infrastructure -- \texttt{I$_3$}   & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99FF99}1
                                     & \cellcolor[HTML]{99AA99}0
                                     & \cellcolor[HTML]{99AA99}0
                                     & \cellcolor[HTML]{99FF99}0 \\
Other -- \texttt{I$_4$}              & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99AA99}3
                                     & \cellcolor[HTML]{99AA99}0
                                     & \cellcolor[HTML]{99FF99}0 \\
\textbf{Total}                       & \cellcolor[HTML]{99FF99}\textbf{3}
                                     & \cellcolor[HTML]{99FF99}\textbf{4}
                                     & \cellcolor[HTML]{99FF99}\textbf{2}
                                     & \cellcolor[HTML]{99AA99}\textbf{6}
                                     & \cellcolor[HTML]{99AA99}\textbf{6}
                                     & \cellcolor[HTML]{99AA99}\textbf{4} \\

@@

                                     & \cellcolor[HTML]{99BB99}\textbf{8} \\

@@

@@

                                     & \cellcolor[HTML]{99FF99}2 \\

@@

@@

                                     & \cellcolor[HTML]{99AA99}\textbf{8} \\

@@

                                     & \cellcolor[HTML]{99AA99}\textbf{23}

@@

>>>>

\todo{add interaction a with the infrastructure for all architectures (application porting)}

@@

\todo{explain that such an abstract architecture has to be concretized to allow for comparison with other architectures. Say also that we could envisage other concretizations but we don't include them as we don't have any experience with these systems.}

@@

\emph{Modularity} measures the ability to replace elements in the
architecture, or to integrate new elements. Elements include
workflow engines, workflows and infrastructure. It breaks down in 4
metrics:
\begin{itemize}[itemsep=0cm]
\item New engine type: number of interactions or components to modify
  to integrate a new workflow engine type in the
  architecture. Modification of an interaction is deemed necessary
  when the parameters involved in this interaction are
  modified. Modification of a component is required when its code
  needs to be modified or recompiled (science gateway or workflow
  service), or when a new piece of software has to be installed
  (infrastructure only). Adding a new type of workflow engine allows
  to execute the workflows that have been described in the language(s)
  used by the new engine.
\item New engine version: number of interactions or components to
  modify to integrate a new version of a workflow engine in the
  architecture, assuming that another version of the same engine type
  is already available. We assume that different versions of a
  workflow engine share the same interface, i.e. they can be invoked
  using the same software. When this is not the case, the different
  versions have to be considered as different engine types. 
\item New infrastructure: number of interactions or components to
  modify to integrate a new type of infrastructure in the
  architecture.
\item New workflow: number of interactions required to integrate a new
  workflow in the architecture.
\item Meta-workflow: ability to describe meta-workflows from existing
  workflows (1: not available; 0: available).
\end{itemize}

@@

\subsection{Nested workflows with service invocation}

@@

\subsection{Workflow import with service invocation}

\paragraph{Integration} Compared to the service architecture, workflow
import requires an additional interaction \texttt{g} in the science
gateway (\texttt{I$_1$}=2, \texttt{I$_2$}=3).

\paragraph{Robustness} Since workflow conversion is not involved in
the execution (it is an offline process), metrics are as in the
service architecture (\texttt{R$_1$=3}, \texttt{R$_2$=4},
\texttt{R$_3$=0}).

\paragraph{Modularity} Since adding a new type of workflow engine aims
at supporting more workflows, we consider that it only requires to
re-implement interaction \texttt{g} in this architecture. Note,
however, that implementing interaction \texttt{g} can require very
substantial work depending on the complexity of the language used by
the new engine (\texttt{M$_1$=1}). Based on the same logic, adding a
new engine version only requires modifying interaction \texttt{g}
(\texttt{M$_2$=1}). As in the service architecture, interfacing with a
new infrastructure requires modifications in the workflow service and
in interactions \texttt{b} and \texttt{c$_2$}
(\texttt{M$_3$=3}). Adding a new workflow is done through interactions
\texttt{a} and \texttt{g} (\texttt{M$_4$=2}). Meta-workflows are
available after import, by connecting workflows in the language used
in the science gateway (\texttt{M$_5$=0}).

\paragraph{Scalability}  Since workflow conversion is not involved in
the execution (it is an offline process), metrics are as in the
service architecture (\texttt{S$_1$=1}, \texttt{S$_2$=1},
\texttt{S$_3$=1}, \texttt{S$_4$=0}).

\begin{table*}
\centering
\begin{tabular}{rcccccc}
\textbf{Metric}                      & \textbf{Tight}
                                     & \textbf{Service}
                                     & \textbf{Sub-task}
                                     & \textbf{Pool}
                                     & \textbf{Nested}
                                     & \textbf{Import} \\
\multicolumn{7}{c}{\cellcolor[HTML]{EEEEEE}\textbf{Integration}}\\
Science gateway -- \texttt{I$_1$}    & \cellcolor[HTML]{99FF99}1
                                     & \cellcolor[HTML]{99FF99}1
                                     & \cellcolor[HTML]{99FF99}0  
                                     & \cellcolor[HTML]{99FF99}1
                                     & \cellcolor[HTML]{99FF99}1
                                     & \cellcolor[HTML]{99AA99}2 \\
Workflow engine -- \texttt{I$_2$}    & \cellcolor[HTML]{99AA99}2
                                     & \cellcolor[HTML]{99FF99}3 
                                     & \cellcolor[HTML]{99FF99}1
                                     & \cellcolor[HTML]{99AA99}2
                                     & \cellcolor[HTML]{99AA99}5
                                     & \cellcolor[HTML]{99FF99}3 \\
Infrastructure -- \texttt{I$_3$}   & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99FF99}1
                                     & \cellcolor[HTML]{99AA99}0
                                     & \cellcolor[HTML]{99AA99}0
                                     & \cellcolor[HTML]{99FF99}0 \\
Other -- \texttt{I$_4$}              & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99AA99}3
                                     & \cellcolor[HTML]{99AA99}0
                                     & \cellcolor[HTML]{99FF99}0 \\
\textbf{Total}                       & \cellcolor[HTML]{99FF99}\textbf{3}
                                     & \cellcolor[HTML]{99FF99}\textbf{4}
                                     & \cellcolor[HTML]{99FF99}\textbf{2}
                                     & \cellcolor[HTML]{99AA99}\textbf{6}
                                     & \cellcolor[HTML]{99AA99}\textbf{6}
                                     & \cellcolor[HTML]{99AA99}\textbf{5} \\

@@

                                     & \cellcolor[HTML]{99BB99}\textbf{7} \\

@@

                                     & \cellcolor[HTML]{99FF99}1 \\

@@

                                     & \cellcolor[HTML]{99FF99}3 \\

@@

                                     & \cellcolor[HTML]{99AA99}0 \\

@@

                                     & \cellcolor[HTML]{99AA99}\textbf{7} \\

@@

                                     & \cellcolor[HTML]{99AA99}\textbf{22}

@@

The difficulty of implementing interaction g is not properly
reflected. For instance, fsl uses bash, which would be very difficult to convert.
